
---
title: "BioSCape Flight Geometry Calculation"
output: github_document
---


```{r, message=FALSE,results='hide'}
library(sf)
library(units)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(ggforce)
library(ggmap)
library(data.table)
library(piggyback)
library(geosphere)
library(smoothr)
source("R/make_flight_line.R")
```


## Define working projection
```{r}
wproj="+proj=utm +zone=34 +south +ellps=WGS84 +datum=WGS84 +units=m +no_defs" # working projection 
```



# Aircraft Details

```{r}
g3_altitude=set_units(c(28000,41000,45000),feet)
g5_altitude=set_units(c(28000,41000,51000),feet)
h_feet=g3_altitude

h=set_units(h_feet, km)  #altitude in km
```



# Instrument Details

## AVIRIS-NG 

Details from https://avirisng.jpl.nasa.gov/specifications.html

```{r}
AVIRISNG=data.frame(
  instrument="AVIRIS-NG",
  row=1, #just used for plotting
  tfov=set_units(36,degrees),
  ifov=set_units(1,mrad),
  spec_n=480,
  spec_res=set_units(5,nm),
  spec_min=set_units(380,nm),
  spec_max=set_units(2510,nm)
)
```

## HyTES 
Details from https://directory.eoportal.org/web/eoportal/airborne-sensors/hytes

```{r}
HyTES=data.frame(
  instrument="HyTES", 
  row=2, #just used for plotting
  tfov=set_units(50,degrees),
  ifov=set_units(1.44,mrad),
  spec_n=256,
  spec_res=set_units(17.6,nm),
  spec_min=set_units(set_units(7.5,um),nm),
  spec_max=set_units(set_units(12,um),nm)
)
```

## PRISM
Details from https://prism.jpl.nasa.gov/spectrometer_char.html

```{r}
PRISM=data.frame(
  instrument="PRISM",
  row=3, #just used for plotting
  tfov=set_units(30.7,degrees),
  ifov=set_units(0.97,mrad),
  spec_n=608,
  spec_res=set_units(2.83,nm),
  spec_min=set_units(349.9,nm),
  spec_max=set_units(1053.5,nm)
)
```

## LVIS
```{r}
LVIS=data.frame(
  instrument="LVIS",
  row=4, #just used for plotting
  tfov=set_units(200,mrad) %>% set_units(degrees), #https://lvis.gsfc.nasa.gov/Data/Maps/ABoVE2019Map.html
  ifov=set_units(0.75,mrad),
  spec_n=1,
  spec_res=set_units(1,nm),
  spec_min=set_units(1064,nm),
  spec_max=set_units(1064,nm)
)
```

## Define angular functions

```{r}
get_ground_fov<- function(h,fov){
  2*set_units(h,km)*tan(set_units(fov/2,radians)) 
}

```

## Instrument Summary

```{r}
sensors=bind_rows(AVIRISNG,HyTES,PRISM,LVIS) %>% 
  rowwise() %>% 
  mutate(
    # get swath widths
    swath_width_low=get_ground_fov(min(h),tfov) %>% set_units(km),
    swath_width_mid=get_ground_fov(median(h),tfov) %>% set_units(km),  
    swath_width_high=get_ground_fov(max(h), tfov) %>% set_units(km),  
    # calculate GIFOV - Ground-projected instantaneous field of view
    gifov_low=get_ground_fov(min(h), ifov)  %>% set_units(m),  
    gifov_mid=get_ground_fov(median(h),ifov) %>% set_units(m),
    gifov_high=get_ground_fov(max(h), ifov)%>% set_units(m)
    ) %>% 
  mutate(
    swath_width_km=paste0("[",round(swath_width_low,2),",",
                          round(swath_width_high,2),"]"),
    gifov_m=paste0("[",round(gifov_low,2),",",
                                round(gifov_high),"]"))

sensors %>% 
  select(instrument,
         spectral_bands=spec_n,spectral_resolution=spec_res,
         swath_width_km,spatial_resolution_m=gifov_m) %>% 
DT::datatable()
```


## Compare with aviris-ng data
Read data from [here](https://avirisng.jpl.nasa.gov/dataportal/)
```{r,message=F, warning=F}

avirisng=googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1UNgRz5EZNLcKtp9cykbi-0lG_D0LWq2lRffwFX3Lpkg/edit#gid=1639195767") %>% 
  filter(grepl("Alt",.$Comments)) # keep only rows that report the altitude

avirisng$alt=set_units(as.numeric(gsub("kft.*$","",gsub("Alt = ","",avirisng$Comments)))*1000,"ft")
# subtract mean scene elevation
avirisng$alt_over_ground=avirisng$alt-set_units(avirisng$`Mean Scene Elevation`,"m")
avirisng$calculated_gifov=get_ground_fov(avirisng$alt_over_ground,set_units(1,mrad)) %>% set_units(m)

avirisng$calculated_gifov_dif=avirisng$calculated_gifov-set_units(avirisng$`Pixel Size`,m)

summary(avirisng$calculated_gifov_dif)


avirisng %>% 
  filter(alt<set_units(70000,ft)) %>% 
  ggplot(aes(x=`Pixel Size`,y=calculated_gifov))+
  geom_hex(bins=100)+
  scale_fill_viridis_c()+
  geom_abline(slope=1,intercept=0,col="red")+
  coord_equal()+
  xlab("Pixel size (m) reported by AVIRIS team")

```

### AVIRIS-ng flight altitudes

```{r}
avirisng %>% 
  filter(alt<set_units(60000,ft)) %>% 
  ggplot(aes(x=alt))+
  geom_histogram()

```

# Typical flights

G3 https://jsc-aircraft-ops.jsc.nasa.gov/gulfstream-giii.html
6 hours / 2,600 nautical miles = 4815.2 km

```{r}
d=4815*(5/6) # 6 hour flight coverage in km (minus one hour for takeoff/landing)
set_units(d,"km")
```


```{r, eval=F, echo=F}
# Flight parameter visualization
altitudes=c(seq(1000,30000,by=10000)) %>% set_units(ft)

sensor_params <- sensors %>% 
  select(instrument,ifov,tfov) %>% 
    expand_grid(altitudes) %>% 
        rowwise() %>% 
        mutate(
          swath_width=get_ground_fov(altitudes,tfov) %>% set_units(km),
          gifov=get_ground_fov(altitudes, ifov)  %>% set_units(m)) %>% 
  gather(key="key",value="value",-instrument,-ifov,-tfov,-altitudes)

p1=sensor_params %>% 
  ggplot(aes(y=altitudes,x=gifov,color=instrument))+
  geom_line()

p2=sensor_params %>% 
  ggplot(aes(y=altitudes,x=swath_width,color=instrument))+
  geom_line()

sensor_params %>% 
  ggplot(aes(y=altitudes,x=value,color=instrument))+
  facet_wrap(~key)+
  geom_line()


plot_grid(p1,p2,byrow = T,align = "v")

```

# Example Path

```{r}
path=st_linestring(x=as.matrix(rbind(
  c(18.46929,-34.38451),
  c(18.37620,-34.08127),
  c(18.46486,-33.99350),
  c(18.77962,-33.94030),
  c(18.93478,-33.84011))),dim="XY") %>% 
  st_sfc() %>% st_sf(path=1,geom=.,crs=4326) %>% 
  smooth(method = "ksmooth")
```

## Calculate instrument polygons
```{r, message=FALSE, output="hide"}
flights <- sensors %>% 
  rowwise() %>% 
  mutate(geometry=make_flight_line(path=path,
                                   swath_width = swath_width_mid)) %>% 
  st_as_sf() %>%
  mutate(instrument=factor(instrument,ordered=T,levels = c("HyTES","AVIRIS-NG","PRISM","LVIS")))


bbox=flights %>% #st_transform(flights,4326) %>% 
  st_buffer(dist = 0.2) %>% 
  st_bbox();names(bbox)=c("left","bottom","right","top")

hdf <- get_map(location=bbox,
              maptype="toner", zoom=12,
              source="stamen")
#hdf <- get_map(location=bbox,  #different background
#               maptype="satellite", zoom=12)
```


## Visualize flight path
```{r}
ggmap(hdf, extent = "normal")+
  geom_sf(data=st_transform(flights,4326),mapping=aes(fill=instrument),
          alpha=.4,color=NA,inherit.aes = F)+
  geom_sf(data=st_transform(path,4326),inherit.aes = F)+
  ylab("Latitude")+
  xlab("Longitude")

```

# Write KML for Google Earth Visualization
```{r}
flights %>% 
  st_segmentize(50) %>% 
#  mutate(OGR_STYLE = "BRUSH(fc:#0000FF80); PEN(c:#FF0000,w:1px)") %>%  #fiddle with kml aesthetics
  st_write("data/transect.kml",append=F,altitudeMode="clampToGround")

if(F){  # if desired, push file to github release
  tag="v0.0.1"
  repo="BioSCape-io/BioSCape_flights"
  pb_new_release(repo, tag)
  pb_upload("data/transect.kml", 
          repo = repo, 
          tag = tag)
}

```